Transmembrane helices are an overlooked and evolutionarily conserved source of major histocompatibility complex class I and II epitopes

### 2021-12-12 frontiers.tex
Use 'citep' instead of 'cite' in LaTeX

The MHC proteins are heterodimeric complexes encoded by the
HLA (Human Leukocyte Antigens) genes.
In humans, the peptide binding groove of MHC-I is made by only the alpha subunit. There are three classical alleles of MHC-I, hallmarked by a highly polymorphic alpha chain called HLA-A, HLA-B and HLA-C, that all present epitopes to cytolytic T cells. 
For MHC-II, both the alpha and the beta chains contribute to the peptide binding groove. There are three classical alleles of MHC-II as well, called HLA-DR, HLA-DQ and HLA-DP, that all present epitopes to helper T cells.
Each MHC complex can present a subset of all possible peptides.
For example, HLA-A and HLA-B have no overlap in which
epitopes they bind \cite{lund2004definition}.
Moreover, the HLA genes of humans are highly polymorphic, with hundreds 
to thousands of different alleles, 
and each different allele presents a different subset of peptides \cite{marsh2010nomenclature}.

@@

Humans express a limited set of MHC alleles
and therefore an individual's immune system detects 
only a fraction of all possible peptide fragments. 
However, at the population level, the coverage of pathogenic peptides that are detected 
is very high, because of the highly polymorphic MHC genes.
It is therefore believed that MHC polymorphism improves immunity at the population level, 
as mutations in a protein that disrupt a particular MHC presentation at the individual level, 
so-called escape mutations, 
will not affect MHC presentation for all alleles present in the population \cite{sommer2005importance}.

@@

Many studies are aimed at identifying the repertoire of epitopes that are presented 
in any of the different alleles to determine which epitopes will result in an immune response, 
as this will for instance aid the design of vaccines. 
These studies have led to the development of prediction algorithms 
that allow for very reliable \emph{in silico} 
predictions of the peptide binding affinities
\cite{larsen2010identification,schellens2008unanticipated,tang2011genome}.
For example, S. Tang et al. \cite{tang2011genome} found that, 
of the 432 peptides that were predicted to bind to an MHC allele,
86\% were experimentally confirmed to do so.

@@

Using these prediction algorithms, 
we recently showed that peptides derived 
from transmembrane helices (TMHs) 
are likely to be more frequently presented by MHC-I 
than expected based on their abundance \cite{bianchi2017},
which is in line with a previous study 
by Istrail et al \cite{istrail2004comparative},
demonstrating that N-terminal signal sequences 
are likely to be presented within major histocompatibility complexes, 
due their hydrophobic nature. 
Moreover, we showed that some well-known immunodominant peptides stem from TMHs. 
This over-presentation is attributed to the fact 
that the peptide-binding groove of most MHC-I alleles
is relatively hydrophobic, 
and therefore hydrophobic TMH-derived peptides have a higher affinity 
to bind than their soluble hydrophobic 
counterparts.

TMHs are hydrophobic 
as they need to span the hydrophobic lipid bilayer of cellular membranes.
They consist of an alpha helix of, on average, 23 amino acids in length. 
TMHs can also be predicted with high accuracy from a protein sequence 
by bioinformatics approaches \cite{krogh2001predicting,kall2004combined,arai2004conpred,jones2007improving,klammer2009metatm,wang2019efficient}. 
For example, a study by Jones \cite{jones2007improving} found that,
from 184 transmembrane proteins (TMPs) with known topology, 
80\% of the TMH predictions of these proteins
matched the experimental findings.
TMHs are common structures in the proteins of humans and microbes. 
Different TMH prediction tools estimate that 15-39\% of all proteins 
in the human proteome contain at least one TMH \cite{ahram2006estimation}.
However, the physiological reason why peptides derived from TMHs 
would be presented more often than peptides 
stemming from soluble (i.e., extracellular or cytoplasmic) protein regions is unknown. 
In this study, we hypothesized that the presentation of 
TMH residues is evolutionarily preferred, 
since TMHs are less prone to undergo escape mutations. 
One reason to expect such a reduced 
variability (and hence evolutionary conservation) in TMHs, 
is that these are restricted in their variability 
by the functional requirement to span a lipid bilayer. 
This limits 
many of the amino acids present in TMHs 
to have hydrophobic side chains 
\cite{hessa2007molecular,jones1994model}.
Therefore, we speculated that the TMHs of pathogens 
might have a lower chance to develop escape mutations, 
as that will result in a dysfunctional TMH 
and render the protein inactive.

This study had two objectives. 
First, we aimed to generalize our findings by predicting
the antigenic presentation
from different kingdoms of life in both MHC-I and -II. 
From these \emph{in silico} predictions, we conclude that TMH-derived
epitopes from a human, viral and bacterial proteome
are likely to be presented more often than expected by chance 
for most alleles of MHC-I and II. 
We confirmed the presentation of TMH-derived peptides 
by re-analysis of peptides from The Immune Epitope Database (IEDB) \cite{vita2019immune}.
Second, we tested our hypothesis that TMHs 
are more evolutionary conserved than soluble protein regions.
Our analysis of human single nucleotide polymorphisms (SNPs) showed 
that random point mutations are indeed less likely
to occur within TMHs. 
These findings strengthen the emerging notion 
that TMHs are important for the T cell-mediated adaptive immune system, 
and hence are of  
overlooked importance in vaccine development.

@@

To predict how frequently epitopes overlapping with TMHs are presented,
a similar analysis strategy was applied as described in \cite{bianchi2017} 
for several alleles of both MHC-I and MHC-II, 
and for a human, viral and bacterial proteome.
To summarize, for each proteome, 
all possible 9-mers (for MHC-I) or 14-mers (MHC-II) were derived. 
For each of these peptides, we determined if it overlapped with a predicted 
TMH and if it was predicted to bind to the most frequent alleles of each MHC allele.

For MHC-I, 9-mers were used, 
as this is the length most frequently presented in MHC-I 
and was used in our earlier study \cite{bianchi2017}. 
For MHC-II, 14-mers were used, 
as this is the most frequently occurring epitope length \cite{bergseng2015different}.
A human (UniProt ID UP000005640\_9606), 
viral (SARS-CoV-2, UniProt ID UP000464024) 
and bacterial (\emph{Mycobacterium tuberculosis}, UniProt ID UP000001584) 
reference proteome was used. 
\verb;TMHMM; \cite{krogh2001predicting} was used to predict the topology 
of the proteins within these proteomes.
To predict the affinity of an epitope to a certain HLA allele,
 \verb;EpitopePrediction; \cite{bianchi2017} for MHC-I 
and \verb;MHCnuggets; \cite{shao2020high} for MHC-II was used.
Both MCH-I and MHC-II alleles were selected 
to have a high prevalence in the population,
where the alleles of MHC-I are the alleles representing the 13 supertypes 
with over 99.6\% coverage of the population's MHC-I repertoire as defined by \cite{lund2004definition} \cite{sette1999},
and the 21 MHC-II alleles, have a phenotypic frequency 
of 14\% or more in the human population \cite{greenbaum2011functional}.

@@

To obtain experimental confirmation that peptides stemming from TMHs 
are presented by MHC-I and MHC-II,
we mined the IEDB \cite{vita2019immune}
for confirmed human MHC-ligands.
We queried the IEDB for all linear epitopes obtained
from MHC ligand assays in healthy humans, 
carrying the MHC alleles as used in this study.
From these epitopes, we kept those that were present
exactly once in the human reference proteome
with UniProt ID UP000005640\_9606.
We predicted the topology of the protein each epitope
was found in, using \verb;TMHMM; \cite{krogh2001predicting},
from which we concluded if the epitope is overlapping with a TMH
with at least 1 amino acid.

@@

As a data source, multiple
NCBI (\url{https://www.ncbi.nlm.nih.gov/}) databases were used: 
the \emph{dbSNP} \cite{sherry2001dbsnp} database,
which contains 650 million 
cataloged non-redundant human variations (called RefSNPs,
\url{https://www.ncbi.nlm.nih.gov/snp/docs/RefSNP_about/}), and the databases \emph{gene} (for gene names \cite{brown2015gene})
and \emph{protein} (for proteins sequences \cite{sayers2010database}).

@@

Figure \ref{fig:1}A shows the predicted presentation of TMH-derived peptides in MHC-I,
for a human, viral and bacterial proteome.
Per MHC-I allele, it shows the percentage of binders that overlap with a TMH 
with at least one residue.
The horizontal line shows the expected percentage of TMH-derived epitopes 
that would be presented, if TMH-derived epitopes would be presented just as 
likely as epitopes derived from soluble regions,
when assuming equal incidence of soluble and TMH-derived epitope presentation.
For 11 out of 13 MHC-I alleles, TMH-derived epitopes are predicted to be presented more often 
than the null expectation, for a human and bacterial proteome.
For the viral proteome, 12 out of 13 MHC-I alleles present
TMH-derived epitopes more often than expected by chance.
The extent of the over-presentation between the different alleles
is similar for the probed proteomes, 
which strengthens our previous conclusion \cite{bianchi2017} 
that the hydrophobicity of the MHC-binding groove 
is the main factor responsible for the predicted over-presentation 
of TMH-derived peptides.

@@

For MHC-I, we previously showed that the over-presentation of TMH-derived 
peptides is caused by the hydrophobicity of the peptide binding 
grooves \cite{bianchi2017}. 
Figures \ref{fig:1}B and \ref{fig:1}C
show the extent of over-presentation
of TMH-derived epitopes as a function of the hydrophobicity preference score 
for the different human MHC alleles.
An assumed linear correlation explains 88\% of the variability in MHC-I.
For MHC-II, 62\% of the variability is explained by hydrophobicity. 
This indicates that TMH-derived peptides are over-presented, 
because the peptide binding grooves of most MHC-I and -II alleles 
are relatively hydrophobic.

@@

Although our data show that 
TMH-derived epitopes are presented in all classical MHC-I and MHC-II alleles, 
the molecular mechanisms of how integral membrane proteins are processed 
for MHC presentation are largely unknown \cite{bianchi2017}. 
Most prominently, the fundamental principles of 
how TMHs are extracted from their hydrophobic lipid environments 
into the aqueous vacuolar lumen, 
leading to subsequent proteolytic processing are unresolved.

A first possibility is that the extraction of TMPs from the membrane 
is mediated by the ER-associated degradation (ERAD) machinery. 
For MHC class I (MHC-I) antigen presentation of soluble proteins, 
the loading of the epitope primarily occurs at the endoplasmatic reticulum (ER). 
The chaperones tapasin (TAPBP), ERp57 (PDIA3), 
and calreticulin (CALR) \cite{rock2016present} first assemble 
and stabilize the heavy and light chains of MHC-I. 
Later, this complex binds to the transporter 
associated with antigen processing (TAP) 
leading to the formation of the so-called peptide-loading complex (PLC). 
The PLC drives import of peptides into the ER 
and mediates their subsequent loading into the peptide-binding groove of MHC-I \cite{blees2017structure}. 
Membrane proteins first will have to be extracted from the membrane 
before they become amenable to this MHC-I loading by the PLC. 
In the ER, this process can be orchestrated by the ERAD machinery, 
consisting of several chaperones that recognize TMPs, 
ubiquitinate them, and extract them from the ER membrane 
into the cytosol (retrotranslocation) for proteasomal degradation \cite{preston2017evolving,meusser2005erad}. 
Similar to the peptides generated from soluble proteins, 
the TMP-derived peptides might then be re-imported by TAP 
into the ER for MHC-I loading. 
This ERAD-driven antigen retrotranslocation might be facilitated by lipid bodies (LBs) \cite{bougneres2009role}, 
since LBs can serve as cytosolic sites for ubiquitination of ER-derived cargo \cite{fujimoto2006proteasomal}.

A second possibility is that TMPs are proteolytically processed 
by intramembrane proteases that cleave TMHs while they are still membrane embedded. Supporting this hypothesis is the well-established notion that peptides generated by signal peptide peptidases (SPPs), an important class of intramembrane proteases that cleave TMH-like signal sequences, 
are presented on a specialized class of MHC-I called HLA-E \cite{oliveira2015alternative}. 
The loading of peptides generated by SPP onto MHC-I does not depend on the proteosome and TAP, 
possibly because the peptides are directly released into the lumen of the ER \cite{oliveira2015alternative}. 
However, this mechanism cannot explain how most membrane proteins can be processed for antigen presentation, because SPPs only cleave TMH-like signal sequences at their C-termini, and N-terminal domains will hence not be removed. 
Nevertheless, the presentation of peptides with a high hydrophobicity index 
was shown to be independent of TAP as well \cite{lautscham2001processing}, 
suggesting that the TMH peptides might perhaps be released directly in the ER lumen by other intramembrane proteases.

A third possibility is that peptide processing and MHC-loading occur in multivesicular bodies (MVBs) \cite{oliveira2015alternative}. 
TMPs can be routed from the plasma membrane and other organelles by vesicular trafficking to endosomes. Eventually, these TMPs can be sorted by the endosomal sorting complexes required for transport (ESCRT) pathway into luminal invaginations that pinch off from the limiting membrane and form intraluminal vesicles. This thus results in MVBs where the membrane proteins destined for degradation are located in intraluminal vesicles. Upon the fusion of MVBs with lysosomes, 
the entire intraluminal vesicles including the TMPs are degraded \cite{gruenberg2020life}. 
Via this mechanism, TMPs might well be processed for antigen presentation, 
particularly since the loading of MHC-II molecules is well understood 
to occur in MVBs \cite{kleijmeer2001reorganization,peters1991segregation,zwart2005spatial}. 
However, such processing of membrane proteins in MVBs for antigen presentation poses a problem, because complexes of HLA-DR with its antigen-loading chaperon HLA-DM were only observed on intraluminal vesicles, 
but not on the limiting membranes of MVBs \cite{zwart2005spatial}, 
indicating that epitope loading of MHC-II also occurs at intraluminal vesicles. This observation hence raises the question how the intraluminal vesicles carrying the TMPs destined for antigen presentation can be selectively degraded, while the intraluminal vesicles carrying the MHC-II remain intact. A second problem is that phagosomes carrying internalized microbes lack intraluminal vesicles, 
and it is hence unclear how TMPs from these microbes 
would be routed to MVBs for MHC-II loading \cite{zwart2005spatial}.

Alternatively to the enzymatic degradation of lipids in MVBs by lipases \cite{sander2016lipase,gilleron2016lysosomal}, 
they might be oxidatively degraded by reactions with radical oxygen species produced by the NADPH oxidase NOX2 \cite{dingjan2016lipid}. 
This oxidation can result in a destabilization and disruption of membranes \cite{dingjan2016lipid} 
and might thereby lead to the extraction of TMPs. 
Due to the hydrophobic nature of TMHs, 
however, the extracted proteins will likely aggregate 
and it is unclear how these aggregates would be processed further for MHC loading.

@@

In general, one might expect that evolutionary selection shapes an immune system 
where surveillance is directed towards protein regions 
essential for the survival, proliferation and/or virulence or pathogenic microbes, 
as these will be most conserved.
In SARS-CoV-2, for example, there is preliminary evidence that the strongest
selection pressure is directed upon residues that change its 
virulence \cite{velazquez2020positive}.
These regions, however, may only account for a small part of a pathogen's proteome.
Additionally, the structure and function of these essential regions might differ widely between different pathogenic proteins.
Because of this scarcity and variance in targets, 
one can imagine that it will be mostly unfeasible 
to provide innate immune responses against such rare essential protein regions, 
as suggested in a study on influenza \cite{han2019individual},
where it was found that the selection pressure
exerted by the immune system was either weak or absent.

@@

There have been more efforts to assess the conservation of TMHs,
using different methodologies.
One such example is a study by Stevens and Arkin \cite{stevens2001substitution}, 
in which aligned protein sequence data was used.
Also this study found that TMHs are evolutionarily more conserved,
as the mean amino acid substitution rate in TMHs is about ten
percent lower,
which is a similar value as we found.
Another example is a study by Oberai, et al. \cite{oberai2009structural} that estimated the conservation
scores for TMHs and soluble regions based on 
alignments of evolutionary related proteins,
and also found that TMHs are more conserved, 
with a conservation score that was 17\% higher in 
TMHs.
Note that the last study also found that mutations in human TMHs are likelier to cause
a disease, in line with our conclusion that TMHs are more conserved.

>>>>

The MHC proteins are heterodimeric complexes encoded by the
HLA (Human Leukocyte Antigens) genes.
In humans, the peptide binding groove of MHC-I is made by only the alpha subunit. There are three classical alleles of MHC-I, hallmarked by a highly polymorphic alpha chain called HLA-A, HLA-B and HLA-C, that all present epitopes to cytolytic T cells. 
For MHC-II, both the alpha and the beta chains contribute to the peptide binding groove. There are three classical alleles of MHC-II as well, called HLA-DR, HLA-DQ and HLA-DP, that all present epitopes to helper T cells.
Each MHC complex can present a subset of all possible peptides.
For example, HLA-A and HLA-B have no overlap in which
epitopes they bind \citep{lund2004definition}.
Moreover, the HLA genes of humans are highly polymorphic, with hundreds 
to thousands of different alleles, 
and each different allele presents a different subset of peptides \citep{marsh2010nomenclature}.

@@

Humans express a limited set of MHC alleles
and therefore an individual's immune system detects 
only a fraction of all possible peptide fragments. 
However, at the population level, the coverage of pathogenic peptides that are detected 
is very high, because of the highly polymorphic MHC genes.
It is therefore believed that MHC polymorphism improves immunity at the population level, 
as mutations in a protein that disrupt a particular MHC presentation at the individual level, 
so-called escape mutations, 
will not affect MHC presentation for all alleles present in the population \citep{sommer2005importance}.

@@

Many studies are aimed at identifying the repertoire of epitopes that are presented 
in any of the different alleles to determine which epitopes will result in an immune response, 
as this will for instance aid the design of vaccines. 
These studies have led to the development of prediction algorithms 
that allow for very reliable \emph{in silico} 
predictions of the peptide binding affinities
\citep{larsen2010identification,schellens2008unanticipated,tang2011genome}.
For example, S. Tang et al. \citep{tang2011genome} found that, 
of the 432 peptides that were predicted to bind to an MHC allele,
86\% were experimentally confirmed to do so.

@@

Using these prediction algorithms, 
we recently showed that peptides derived 
from transmembrane helices (TMHs) 
are likely to be more frequently presented by MHC-I 
than expected based on their abundance \citep{bianchi2017},
which is in line with a previous study 
by Istrail et al \citep{istrail2004comparative},
demonstrating that N-terminal signal sequences 
are likely to be presented within major histocompatibility complexes, 
due their hydrophobic nature. 
Moreover, we showed that some well-known immunodominant peptides stem from TMHs. 
This over-presentation is attributed to the fact 
that the peptide-binding groove of most MHC-I alleles
is relatively hydrophobic, 
and therefore hydrophobic TMH-derived peptides have a higher affinity 
to bind than their soluble hydrophobic 
counterparts.

TMHs are hydrophobic 
as they need to span the hydrophobic lipid bilayer of cellular membranes.
They consist of an alpha helix of, on average, 23 amino acids in length. 
TMHs can also be predicted with high accuracy from a protein sequence 
by bioinformatics approaches \citep{krogh2001predicting,kall2004combined,arai2004conpred,jones2007improving,klammer2009metatm,wang2019efficient}. 
For example, a study by Jones \citep{jones2007improving} found that,
from 184 transmembrane proteins (TMPs) with known topology, 
80\% of the TMH predictions of these proteins
matched the experimental findings.
TMHs are common structures in the proteins of humans and microbes. 
Different TMH prediction tools estimate that 15-39\% of all proteins 
in the human proteome contain at least one TMH \citep{ahram2006estimation}.
However, the physiological reason why peptides derived from TMHs 
would be presented more often than peptides 
stemming from soluble (i.e., extracellular or cytoplasmic) protein regions is unknown. 
In this study, we hypothesized that the presentation of 
TMH residues is evolutionarily preferred, 
since TMHs are less prone to undergo escape mutations. 
One reason to expect such a reduced 
variability (and hence evolutionary conservation) in TMHs, 
is that these are restricted in their variability 
by the functional requirement to span a lipid bilayer. 
This limits 
many of the amino acids present in TMHs 
to have hydrophobic side chains 
\citep{hessa2007molecular,jones1994model}.
Therefore, we speculated that the TMHs of pathogens 
might have a lower chance to develop escape mutations, 
as that will result in a dysfunctional TMH 
and render the protein inactive.

This study had two objectives. 
First, we aimed to generalize our findings by predicting
the antigenic presentation
from different kingdoms of life in both MHC-I and -II. 
From these \emph{in silico} predictions, we conclude that TMH-derived
epitopes from a human, viral and bacterial proteome
are likely to be presented more often than expected by chance 
for most alleles of MHC-I and II. 
We confirmed the presentation of TMH-derived peptides 
by re-analysis of peptides from The Immune Epitope Database (IEDB) \citep{vita2019immune}.
Second, we tested our hypothesis that TMHs 
are more evolutionary conserved than soluble protein regions.
Our analysis of human single nucleotide polymorphisms (SNPs) showed 
that random point mutations are indeed less likely
to occur within TMHs. 
These findings strengthen the emerging notion 
that TMHs are important for the T cell-mediated adaptive immune system, 
and hence are of  
overlooked importance in vaccine development.

@@

To predict how frequently epitopes overlapping with TMHs are presented,
a similar analysis strategy was applied as described in \citep{bianchi2017} 
for several alleles of both MHC-I and MHC-II, 
and for a human, viral and bacterial proteome.
To summarize, for each proteome, 
all possible 9-mers (for MHC-I) or 14-mers (MHC-II) were derived. 
For each of these peptides, we determined if it overlapped with a predicted 
TMH and if it was predicted to bind to the most frequent alleles of each MHC allele.

For MHC-I, 9-mers were used, 
as this is the length most frequently presented in MHC-I 
and was used in our earlier study \citep{bianchi2017}. 
For MHC-II, 14-mers were used, 
as this is the most frequently occurring epitope length \citep{bergseng2015different}.
A human (UniProt ID UP000005640\_9606), 
viral (SARS-CoV-2, UniProt ID UP000464024) 
and bacterial (\emph{Mycobacterium tuberculosis}, UniProt ID UP000001584) 
reference proteome was used. 
\verb;TMHMM; \citep{krogh2001predicting} was used to predict the topology 
of the proteins within these proteomes.
To predict the affinity of an epitope to a certain HLA allele,
 \verb;EpitopePrediction; \citep{bianchi2017} for MHC-I 
and \verb;MHCnuggets; \citep{shao2020high} for MHC-II was used.
Both MCH-I and MHC-II alleles were selected 
to have a high prevalence in the population,
where the alleles of MHC-I are the alleles representing the 13 supertypes 
with over 99.6\% coverage of the population's MHC-I repertoire as defined by \citep{lund2004definition} \citep{sette1999},
and the 21 MHC-II alleles, have a phenotypic frequency 
of 14\% or more in the human population \citep{greenbaum2011functional}.

@@

To obtain experimental confirmation that peptides stemming from TMHs 
are presented by MHC-I and MHC-II,
we mined the IEDB \citep{vita2019immune}
for confirmed human MHC-ligands.
We queried the IEDB for all linear epitopes obtained
from MHC ligand assays in healthy humans, 
carrying the MHC alleles as used in this study.
From these epitopes, we kept those that were present
exactly once in the human reference proteome
with UniProt ID UP000005640\_9606.
We predicted the topology of the protein each epitope
was found in, using \verb;TMHMM; \citep{krogh2001predicting},
from which we concluded if the epitope is overlapping with a TMH
with at least 1 amino acid.

@@

As a data source, multiple
NCBI (\url{https://www.ncbi.nlm.nih.gov/}) databases were used: 
the \emph{dbSNP} \citep{sherry2001dbsnp} database,
which contains 650 million 
cataloged non-redundant human variations (called RefSNPs,
\url{https://www.ncbi.nlm.nih.gov/snp/docs/RefSNP_about/}), and the databases \emph{gene} (for gene names \citep{brown2015gene})
and \emph{protein} (for proteins sequences \citep{sayers2010database}).

@@

Figure \ref{fig:1}A shows the predicted presentation of TMH-derived peptides in MHC-I,
for a human, viral and bacterial proteome.
Per MHC-I allele, it shows the percentage of binders that overlap with a TMH 
with at least one residue.
The horizontal line shows the expected percentage of TMH-derived epitopes 
that would be presented, if TMH-derived epitopes would be presented just as 
likely as epitopes derived from soluble regions,
when assuming equal incidence of soluble and TMH-derived epitope presentation.
For 11 out of 13 MHC-I alleles, TMH-derived epitopes are predicted to be presented more often 
than the null expectation, for a human and bacterial proteome.
For the viral proteome, 12 out of 13 MHC-I alleles present
TMH-derived epitopes more often than expected by chance.
The extent of the over-presentation between the different alleles
is similar for the probed proteomes, 
which strengthens our previous conclusion \citep{bianchi2017} 
that the hydrophobicity of the MHC-binding groove 
is the main factor responsible for the predicted over-presentation 
of TMH-derived peptides.

@@

For MHC-I, we previously showed that the over-presentation of TMH-derived 
peptides is caused by the hydrophobicity of the peptide binding 
grooves \citep{bianchi2017}. 
Figures \ref{fig:1}B and \ref{fig:1}C
show the extent of over-presentation
of TMH-derived epitopes as a function of the hydrophobicity preference score 
for the different human MHC alleles.
An assumed linear correlation explains 88\% of the variability in MHC-I.
For MHC-II, 62\% of the variability is explained by hydrophobicity. 
This indicates that TMH-derived peptides are over-presented, 
because the peptide binding grooves of most MHC-I and -II alleles 
are relatively hydrophobic.

@@

Although our data show that 
TMH-derived epitopes are presented in all classical MHC-I and MHC-II alleles, 
the molecular mechanisms of how integral membrane proteins are processed 
for MHC presentation are largely unknown \citep{bianchi2017}. 
Most prominently, the fundamental principles of 
how TMHs are extracted from their hydrophobic lipid environments 
into the aqueous vacuolar lumen, 
leading to subsequent proteolytic processing are unresolved.

A first possibility is that the extraction of TMPs from the membrane 
is mediated by the ER-associated degradation (ERAD) machinery. 
For MHC class I (MHC-I) antigen presentation of soluble proteins, 
the loading of the epitope primarily occurs at the endoplasmatic reticulum (ER). 
The chaperones tapasin (TAPBP), ERp57 (PDIA3), 
and calreticulin (CALR) \citep{rock2016present} first assemble 
and stabilize the heavy and light chains of MHC-I. 
Later, this complex binds to the transporter 
associated with antigen processing (TAP) 
leading to the formation of the so-called peptide-loading complex (PLC). 
The PLC drives import of peptides into the ER 
and mediates their subsequent loading into the peptide-binding groove of MHC-I \citep{blees2017structure}. 
Membrane proteins first will have to be extracted from the membrane 
before they become amenable to this MHC-I loading by the PLC. 
In the ER, this process can be orchestrated by the ERAD machinery, 
consisting of several chaperones that recognize TMPs, 
ubiquitinate them, and extract them from the ER membrane 
into the cytosol (retrotranslocation) for proteasomal degradation \citep{preston2017evolving,meusser2005erad}. 
Similar to the peptides generated from soluble proteins, 
the TMP-derived peptides might then be re-imported by TAP 
into the ER for MHC-I loading. 
This ERAD-driven antigen retrotranslocation might be facilitated by lipid bodies (LBs) \citep{bougneres2009role}, 
since LBs can serve as cytosolic sites for ubiquitination of ER-derived cargo \citep{fujimoto2006proteasomal}.

A second possibility is that TMPs are proteolytically processed 
by intramembrane proteases that cleave TMHs while they are still membrane embedded. Supporting this hypothesis is the well-established notion that peptides generated by signal peptide peptidases (SPPs), an important class of intramembrane proteases that cleave TMH-like signal sequences, 
are presented on a specialized class of MHC-I called HLA-E \citep{oliveira2015alternative}. 
The loading of peptides generated by SPP onto MHC-I does not depend on the proteosome and TAP, 
possibly because the peptides are directly released into the lumen of the ER \citep{oliveira2015alternative}. 
However, this mechanism cannot explain how most membrane proteins can be processed for antigen presentation, because SPPs only cleave TMH-like signal sequences at their C-termini, and N-terminal domains will hence not be removed. 
Nevertheless, the presentation of peptides with a high hydrophobicity index 
was shown to be independent of TAP as well \citep{lautscham2001processing}, 
suggesting that the TMH peptides might perhaps be released directly in the ER lumen by other intramembrane proteases.

A third possibility is that peptide processing and MHC-loading occur in multivesicular bodies (MVBs) \citep{oliveira2015alternative}. 
TMPs can be routed from the plasma membrane and other organelles by vesicular trafficking to endosomes. Eventually, these TMPs can be sorted by the endosomal sorting complexes required for transport (ESCRT) pathway into luminal invaginations that pinch off from the limiting membrane and form intraluminal vesicles. This thus results in MVBs where the membrane proteins destined for degradation are located in intraluminal vesicles. Upon the fusion of MVBs with lysosomes, 
the entire intraluminal vesicles including the TMPs are degraded \citep{gruenberg2020life}. 
Via this mechanism, TMPs might well be processed for antigen presentation, 
particularly since the loading of MHC-II molecules is well understood 
to occur in MVBs \citep{kleijmeer2001reorganization,peters1991segregation,zwart2005spatial}. 
However, such processing of membrane proteins in MVBs for antigen presentation poses a problem, because complexes of HLA-DR with its antigen-loading chaperon HLA-DM were only observed on intraluminal vesicles, 
but not on the limiting membranes of MVBs \citep{zwart2005spatial}, 
indicating that epitope loading of MHC-II also occurs at intraluminal vesicles. This observation hence raises the question how the intraluminal vesicles carrying the TMPs destined for antigen presentation can be selectively degraded, while the intraluminal vesicles carrying the MHC-II remain intact. A second problem is that phagosomes carrying internalized microbes lack intraluminal vesicles, 
and it is hence unclear how TMPs from these microbes 
would be routed to MVBs for MHC-II loading \citep{zwart2005spatial}.

Alternatively to the enzymatic degradation of lipids in MVBs by lipases \citep{sander2016lipase,gilleron2016lysosomal}, 
they might be oxidatively degraded by reactions with radical oxygen species produced by the NADPH oxidase NOX2 \citep{dingjan2016lipid}. 
This oxidation can result in a destabilization and disruption of membranes \citep{dingjan2016lipid} 
and might thereby lead to the extraction of TMPs. 
Due to the hydrophobic nature of TMHs, 
however, the extracted proteins will likely aggregate 
and it is unclear how these aggregates would be processed further for MHC loading.

@@

In general, one might expect that evolutionary selection shapes an immune system 
where surveillance is directed towards protein regions 
essential for the survival, proliferation and/or virulence or pathogenic microbes, 
as these will be most conserved.
In SARS-CoV-2, for example, there is preliminary evidence that the strongest
selection pressure is directed upon residues that change its 
virulence \citep{velazquez2020positive}.
These regions, however, may only account for a small part of a pathogen's proteome.
Additionally, the structure and function of these essential regions might differ widely between different pathogenic proteins.
Because of this scarcity and variance in targets, 
one can imagine that it will be mostly unfeasible 
to provide innate immune responses against such rare essential protein regions, 
as suggested in a study on influenza \citep{han2019individual},
where it was found that the selection pressure
exerted by the immune system was either weak or absent.

@@

There have been more efforts to assess the conservation of TMHs,
using different methodologies.
One such example is a study by Stevens and Arkin \citep{stevens2001substitution}, 
in which aligned protein sequence data was used.
Also this study found that TMHs are evolutionarily more conserved,
as the mean amino acid substitution rate in TMHs is about ten
percent lower,
which is a similar value as we found.
Another example is a study by Oberai, et al. \citep{oberai2009structural} that estimated the conservation
scores for TMHs and soluble regions based on 
alignments of evolutionary related proteins,
and also found that TMHs are more conserved, 
with a conservation score that was 17\% higher in 
TMHs.
Note that the last study also found that mutations in human TMHs are likelier to cause
a disease, in line with our conclusion that TMHs are more conserved.
